\pkg{DNest4}: Diffusive Nested Sampling in \proglang{C++} and \proglang{Python}

### 2016-10-17 paper/ms.tex
Re-running

In Mac OS X, the final line might need to provide information about
your OS version, like so:

\begin{CodeChunk}
\begin{CodeInput}
> MACOSX_DEPLOYMENT_TARGET=10.9 python setup.py install
\end{CodeInput}
\end{CodeChunk}

@@

\section[Running DNest4]{Running \pkg{DNest4}}\label{sec:running}
To demonstrate \pkg{DNest4},
we will use a simple linear regression example where the
sampling distribution is
\begin{align}
y_i | m, b, \sigma &\sim \textnormal{Normal}(mx_i + b, \sigma^2)
\end{align}
and the priors are
\begin{align}
m &\sim \textnormal{Normal}(0, 1000^2)\\
b &\sim \textnormal{Normal}(0, 1000^2)\\
\ln\sigma &\sim \textnormal{Uniform}(-10, 10).
\end{align}
These are naive diffuse priors and do not have any special status.
The dataset is shown in Figure~\ref{fig:regression_lines}
and the code is included in the {\tt code/Examples/StraightLine}
subdirectory, and a slightly simplified version of the
code is explained in Section~\ref{sec:models}.
To execute \pkg{DNest4} on this problem, go to this directory and
execute {\tt main}. The output to the screen
should contain information about levels and ``saving particles to disk''.
After 5000 particles have been saved to the disk, the run will terminate.

@@

\begin{CodeChunk}
\begin{CodeInput}
import dnest4
dnest4.postprocess()
\end{CodeInput}
\end{CodeChunk}
This produces the three diagnostic plots in Figures~\ref{fig:fig1},
\ref{fig:fig2}, and~\ref{fig:fig3}, along with the following output:
\begin{CodeChunk}
\begin{CodeOutput}
log(Z) = -175.545067073
Information = 15.2168377939 nats.
Effective sample size = 658.958278958
\end{CodeOutput}
\end{CodeChunk}
These are the natural log of the marginal likelihood, the
information
\begin{align}
\mathcal{H} &= \int p(\theta|D, M)
\ln\left[\frac{p(\theta | D, M)}{p(\theta | M)}\right] \, d\theta,
\end{align}
which quantifies the degree to which $D$ restricted the
range of possible $\theta$ values,
and the effective sample size,
or number of saved particles with significant posterior weight.
The \code{postprocess} function also saves a file,
{\tt posterior\_sample.txt}, containing posterior samples (one per row).
Unfortunately, it is harder to compute justified error bars on $\ln(Z)$
in DNS than it is in standard NS.
%The \code{postprocess} function does include some options
%for calculating error bars, but they are not used by default,
%and we do not describe them here.
%Hence, the zero-sized error bars on $\ln(Z)$ and $\mathcal{H}$ should
%be ignored.

@@

\begin{CodeChunk}
\begin{CodeInput}
# File containing parameters for DNest4
# Put comments at the top, or at the end of the line.
5       # Number of particles
10000   # new level interval
10000   # save interval
100     # threadSteps - pooling interval
0       # maximum number of levels (0 ==> automatic)
10      # Backtracking scale length (lambda in the paper)
100     # Equal weight enforcement. Beta in the paper
5000    # Maximum number of saves (0 ==> run forever)
\end{CodeInput}
\end{CodeChunk}

>>>>

In Mac OS X, the final line (which installs the \proglang{Python} parts)
of \pkg{DNest4})
needs to provide information about
your OS version. For example, if your computer runs Mac OS X 10.9,
the installation command for the \proglang{Python} package is
\begin{CodeChunk}
\begin{CodeInput}
> MACOSX_DEPLOYMENT_TARGET=10.9 python setup.py install
\end{CodeInput}
\end{CodeChunk}

@@

\section[Running DNest4]{Running \pkg{DNest4}}\label{sec:running}
To demonstrate \pkg{DNest4},
we will use a simple linear regression example where the
sampling distribution is
\begin{align}
y_i | m, b, \sigma &\sim \textnormal{Normal}(mx_i + b, \sigma^2)
\end{align}
and the priors are
\begin{align}
m &\sim \textnormal{Normal}(0, 1000^2)\\
b &\sim \textnormal{Normal}(0, 1000^2)\\
\ln\sigma &\sim \textnormal{Uniform}(-10, 10).
\end{align}
These are naive diffuse priors and do not have any special status.
The dataset is shown in Figure~\ref{fig:regression_lines}
and the code is included in the {\tt code/Examples/StraightLine}
subdirectory, and a slightly simplified version of the
code is explained in Section~\ref{sec:models}.
To execute \pkg{DNest4} on this problem, go to this directory and
execute {\tt main}. The output to the screen
should contain information about levels and ``saving particles to disk''.
After 10,000 particles have been saved to the disk, the run will terminate.

@@

\begin{CodeChunk}
\begin{CodeInput}
import dnest4
dnest4.postprocess()
\end{CodeInput}
\end{CodeChunk}
This produces the three diagnostic plots in Figures~\ref{fig:fig1},
\ref{fig:fig2}, and~\ref{fig:fig3}, along with the following output:
\begin{CodeChunk}
\begin{CodeOutput}
log(Z) = -175.512920079
Information = 15.1683473499 nats.
Effective sample size = 1521.58236273
\end{CodeOutput}
\end{CodeChunk}
These are the natural log of the marginal likelihood, the
information
\begin{align}
\mathcal{H} &= \int p(\theta|D, M)
\ln\left[\frac{p(\theta | D, M)}{p(\theta | M)}\right] \, d\theta,
\end{align}
which quantifies the degree to which $D$ restricted the
range of possible $\theta$ values,
and the effective sample size,
or number of saved particles with significant posterior weight.
The \code{postprocess} function also saves a file,
{\tt posterior\_sample.txt}, containing posterior samples (one per row).
Unfortunately, it is harder to compute justified error bars on $\ln(Z)$
in DNS than it is in standard NS.
%The \code{postprocess} function does include some options
%for calculating error bars, but they are not used by default,
%and we do not describe them here.
%Hence, the zero-sized error bars on $\ln(Z)$ and $\mathcal{H}$ should
%be ignored.

@@

\begin{CodeChunk}
\begin{CodeInput}
# File containing parameters for DNest4
# Put comments at the top, or at the end of the line.
5       # Number of particles
10000   # new level interval
10000   # save interval
100     # threadSteps - pooling interval
0       # maximum number of levels (0 ==> automatic)
10      # Backtracking scale length (lambda in the paper)
100     # Equal weight enforcement. Beta in the paper
10000   # Maximum number of saves (0 ==> run forever)
\end{CodeInput}
\end{CodeChunk}
